% Author: John H. Johnson
% email: john.j.jr@gmail.com
%
% Brief description of file
% -------------------------
%
% This is a chapter in my dissertation tentatively entitled "A New and
% Simpler Central Sets Theorem".  The new and simpler part is due to a
% different definition of J-sets, which in turns leads to a different
% definition of C-sets.  Happily the new definition is equivalent to
% the old definition.  Therefore no proofs need to be rewritten; but
% there is value in rewriting the proofs, since as this chapter shows,
% the rewritten proofs are often simpler than the old proofs.
%
% N.B. The new definition of J-sets essentially first appears in an
% article, "Ramsey Theory in Noncommutative Semigroups", by Bergelson
% and Hindman in their Theorem 2.6.

\documentclass[12pt,showtrims]{memoir}

\usepackage{amsthm, amssymb, amsmath}
\usepackage{color}
\usepackage{endnotes}
\usepackage{url}


\usepackage[margin=1in]{geometry}


\newtheoremstyle{plain}{3mm}{3mm}{\slshape}{}{\bfseries}{.}{.5em}{}
\theoremstyle{plain}

% Numbered theorems 
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{up}[thm]{Ultrafilter Principle}
\newtheorem{radoSelect}[thm]{Rado's Selection Lemma}
\newtheorem{php}[thm]{Pigeonhole Principle}
\newtheorem{vdw}[thm]{Van der Waerden's Theorem}
\newtheorem{hj}[thm]{Hales-Jewett Theorem}
\newtheorem{shj}[thm]{Special Hales-Jewett Theorem}


\newtheorem{FST}[thm]{Hindman's Theorem}
\newtheorem{MBR}[thm]{Multiple Birkhoff Recurrence Theorem}
\newtheorem{recur}[thm]{Recurrence Theorem}
\newtheorem{OCST}[thm]{Furstenburg's Original Central Sets Theorem}
\newtheorem{cst}[thm]{Central Sets Theorem}



\newtheorem{claim}[thm]{Claim}
\newtheorem{ques}[thm]{Question}
\newtheorem{conj}[thm]{Conjecture}


\theoremstyle{definition}

% Numbered "definition" style theorem environments
\newtheorem{defn}[thm]{Definition}
\newtheorem{rmk}[thm]{Remark}
\newtheorem{example}[thm]{Example}

\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\bbN}{\mathbb{N}}
\newcommand{\bbZ}{\mathbb{Z}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\AP}{\mathcal{AP}}
\newcommand{\AL}{\mathcal{AL}}

% Short names for calligraphic math letters.
\newcommand{\calA}{\mathcal{A}}
\newcommand{\calB}{\mathcal{B}}
\newcommand{\calC}{\mathcal{C}}
\newcommand{\calE}{\mathcal{E}}
\newcommand{\calF}{\mathcal{F}}
\newcommand{\calG}{\mathcal{G}}
\newcommand{\calH}{\mathcal{H}}
\newcommand{\calI}{\mathcal{I}}
\newcommand{\calJ}{\mathcal{J}}
\newcommand{\calP}{\mathcal{P}}
\newcommand{\calR}{\mathcal{R}}
\newcommand{\calS}{\mathcal{S}}
\newcommand{\calT}{\mathcal{T}}
\newcommand{\calU}{\mathcal{U}}

\newcommand{\Pf}{\mathcal{P}_f}


\newcommand{\setfunc}[2]{\hbox{${}^{\hbox{$#1$}}\hskip -1 pt #2$}}

\font\bigmath=cmsy10 scaled \magstep 3
\newcommand{\bigtimes}{\hbox{\bigmath \char'2}}

\newcommand{\cchi}{\raise 2 pt \hbox{$\chi$}}

\begin{document}
\addtolength{\baselineskip}{1.7pt}
\section{Hales-Jewett Theorem}
In the article \cite{Hales:1963fk}, Hales and Jewett abstracted van der Waerden's combinatorial argument into a powerful theorem which implies (among other things) van der Waerden's Theorem itself. 
The vast applicability of the Hales-Jewett Theorem lies in the fact that it relates finite coloring of free semigroups and `combinatorial lines'. 
(Combinatorial lines are an abstraction of arithmetic progressions.)
In this chapter, one of our first tasks shall be to import the notion of combinatorial lines into other types of semigroups. 
However before we can get started we must first state some terminology related to the Hales-Jewett Theorem itself.

\begin{defn}
  Let $A$ be a nonempty finite set, let $\star$ denote an element not in $A$, and let $S$ be the free semigroup generated by $A$.
  \begin{itemize}
    \item[(a)] 
      A \emph{variable word (on $A$)} $w(\star)$ is a word in the  free semigroup on the alphabet $A \cup \{\star\}$ in which $\star$ occurs.

    \item[(b)] 
      Given a variable word $w(\star)$ on $A$ and $a \in A$, we let $w(a)$ denote the word in $S$ where each occurrence of $\star$ in $w(\star)$ is replaced by $a$.

    \item[(c)] 
      Given a variable word $w(\star)$ on $A$ we call the set $\{\, w(a) : a \in A \,\}$ a \emph{combinatorial line}.
  \end{itemize}
\end{defn}

\begin{hj}
  Let $A$ be a nonempty finite set, let $S$ be the free semigroup generated by $A$, and let $r$ be a positive integer.
  If $S = \bigcup_{i=1}^r C_i$, then there exists $i \in \{1, 2, \ldots, r\}$ such that $C_i$ contains a combinatorial line, that is, there exists a variable word $w(\star)$ with $\bigl\{\, w(a) : a \in A \,\bigr\} \subseteq C_i$.
\end{hj}

To import the notion of a combinatorial line into other semigroups we shall specialized the notion of a combinatorial line.
The following theorem shows that the conclusion of the Hales-Jewett Theorem may be strengthen. 

\begin{thm}
  \label{thm:special-hj}
  Let $r$ be a positive integer and $A$ a nonempty finite set.
  If $S$ is the free semigroup generated by $A$ and $S = \bigcup_{i=1}^r C_i$, then there exist $i \in \{1, 2, \ldots, r\}$ and a variable word $w(\star)$ which begins and ends with a constant letter and has no two adjacent variable letters such that $\bigl\{\, w(a) : a \in A \,\bigr\} \subseteq C_i$.
\end{thm}
\begin{proof}
  Define the function $\varphi \colon S \to \{1, 2, \ldots, r\}$ by $\varphi(w) = \min\{i : w \in C_i\}$. 
  Fix $b \in A$ and define $\psi \colon S \to S$ as follows: if $w = a_1a_2 \cdots a_n$ with each $a_i \in A$, then $\psi(w) = ba_1ba_2b\cdots a_nb$.
  We have that $\varphi \circ \psi \colon S \to \{1, 2, \ldots, r\}$.
  By the Hales-Jewett Theorem, pick $i \in \{1, 2, \ldots, r\}$ and a variable word $w(\star)$ such that $\{\, w(a) : a \in A \,\} \subseteq (\varphi \circ \psi)^{-1}[i]$.
  Observe that $\psi\bigl(w(\star)\bigr)$ is a variable word that begins and ends with a constant, doesn't have any adjacent variables, and $\{\, \psi\bigl(w(a)\bigr) : a \in A \} \subseteq C_i$.
\end{proof}

\begin{thm}
  \label{thm:finitary-shj}
  Let $k$ be a positive integer.
  We consider $\{1, 2, \ldots, k\}$ as a finite alphabet on $k$ symbols, and for every $n \in \bbN$, we consider $\{1, 2, \ldots, k\}^n$ as the set of all length $n$ words generated by $\{1, 2, \ldots, k\}$.

  Then for each $k$ and $r$ in $\bbN$ there exists $n \in \bbN$ such that whenever $\{1, 2, \ldots, k\}^n$ is $r$-colored, that is, $\varphi \colon \{1, 2, \ldots, k\}^n \to \{1, 2, \ldots, r\}$, there exists a variable word $w(\star)$ which begins and ends with a constant letter and has no two adjacent variable letters such that $\bigl\{\, w(i) : i \in \{1, 2, \ldots, k\} \,\bigr\}$ is monochromatic, that is, $\varphi$ is constant on $\bigl\{\, w(i) : i \in \{1, 2, \ldots, k\} \,\bigr\}$.
\end{thm}
\begin{proof}
  Let $k$ and $r$ be given and suppose that the conclusion fails.
  That is for each $n \in \bbN$ choose an $r$-coloring $\varphi_n \colon \{1, 2, \ldots, k\}^n \to \{1, 2, \ldots, r\}$ such that no variable word which begins and ends with a constant and has no two adjacent variable letters is monochromatic for $\varphi_n$.
  
  Put $S = \bigcup_{n=1}^\infty \{1, 2, \ldots, k\}^n$ and define $\varphi \colon S \to \{1, 2, \ldots, r\}$ as follows: If $w$ is a length $n$ word with $w = (x_1, x_2, \ldots, x_n)$ and each $x_i \in \{1, 2, \ldots, k\}$, then $\varphi(w) = \varphi_n(w)$.
  By Theorem \ref{thm:special-hj} pick $i \in \{1, 2, \ldots, r\}$ and a variable word $w(\star)$ which beings and ends with a constant letter and has no two adjacent variable letters with $\bigl\{\, w(i) : i \in \{1, 2, \ldots, k\} \,\bigr\} \subseteq \varphi^{-1}[\{i\}]$.
  Suppose that the length of $w(\star)$ is $n$, then, $\varphi_n$ is constant on $\bigl\{\, w(i) : i \in \{1, 2, \ldots, k\} \,\bigr\}$, a contradiction. 
\end{proof}

\section{$J$-sets}
The way we shall import the notion of a combinatorial line into other semigroups is by defining the concept of a $J$-set. 

\begin{defn}
  Let $(S, \cdot)$ be a semigroup.
  \begin{itemize}
    \item[(b)] 
      For each $m \in \bbN$, define
      \[
        \calJ_m = \{\, (t_1, t_2, \ldots, t_m) \in \bbN^m : t_1 < t_2 < \cdots < t_m \,\}.
      \]

    \item[(b)] 
      Put $\calT(S) = \setfunc{\bbN}{S}$.
      If the semigroup is clear from context, we write $\calT$ instead of $\calT(S)$.

    \item[(c)] 
      For each $m \in \bbN$, $a \in S^{m+1}$, $t \in \calJ_m$, and $f \in \calT$, put
      \[
        \textstyle
        x(m, a, t, f) = \Bigl( \prod_{i=1}^m \bigl( a(i)f(t_i) \bigr) \Bigr)a(m+1).
      \]

    \item[(d)] 
      We call a subset $A \subseteq S$ a \emph{$J$-set (in $S$)} if and only if for every $F \in \Pf(\calT)$ there exist $m \in \bbN$, $a \in S^{m+1}$, and $t \in \calJ_m$ such that for each $f \in F$ we have $x(m, a, t, f) \in A$.
  \end{itemize}
\end{defn}
\begin{rmk}
  Despite appearances to the contrary, I must point out that $J$-sets are \textsl{not} named after the author!
  In particular, the term $J$-set is derived from the term $J_Y$ set which first appeared as a definition in \cite[Definition 2.4(b)]{Hindman:1996fk}.
\end{rmk}

Our definition of a $J$-set looks different from the definitions given in \cite[Definition 2.2(a)]{Hindman:2009vn} and \cite[Definition 3.3(d)]{De:2008uq}.
The first result we prove about $J$-sets shall be that all of these definitions are equivalent.
However before proving this result we introduce some temporary notation that is analogous to $\calJ_m$.

\begin{defn}
  For each $m \in \bbN$ define
  \begin{align*}
    \calI_m &= \Bigl\{\, \bigl( H(1), H(2), \ldots, H(m) \bigr) : \hbox{$H(i) \in \mathcal{P}_f(\bbN)$ for all $i \in \{1, 2, \ldots, m\}$} \\
        &\hspace{3em} \hbox{ and $\max H(i) < \min H(i+1)$ for all $i \in \{1, 2, \ldots, m-1\}$} \Bigr\}.
  \end{align*}
\end{defn}

Given a $A$ a subset of some semigroup $S$, in \cite[Definition 2.2(a)]{Hindman:2009vn} $A$ is considered a $J$-set if and only if for every $F \in \calT$, there exists $m \in \bbN$, $a \in S^{m+1}$, and $H \in \calI_m$ such that for all $f \in F$, $\prod_{i=1}^n\bigl( a(i) \prod_{t \in H(i)} f(t) \bigr) a(m+1) \in A$.
To show that our definition of a $J$-set is equivalent to \cite[Definition 2.2(a)]{Hindman:2009vn} we will use the following technical `rewriting' lemma.

\begin{lem}
  \label{lem:rewrite-jset}
  Let $(S, \cdot)$ be a semigroup, $m \in \bbN$, $a \in S^{m+1}$, $H \in \calI_m$, and $F \in \Pf(\calT)$. 
  Fix $b \in S$ and for each $f \in F$ define $g_f \in \calT$ by $g_f(t) = f(t)b$. 
  Then there exists $n \in \bbN$, $c \in S^{n+1}$, and $t \in \calJ_n$ such that for all $f \in F$, $x(n, c, t, f) = \prod_{i=1}^m \bigl( a(i) \prod_{t \in H(i)} g_f(t)\big) a(m+1)$.
\end{lem}
\begin{proof}
  Put $H(0) = \emptyset$ and for each $s \in \{0, 1, \ldots, m\}$ define $h_s = \sum_{i=0}^s |H(i)|$.
  Put $n = h_m$ and enumerate $\bigcup_{i=1}^m H(i)$ as a strictly increasing sequence $t_1 < t_2 < \cdots < t_n$ in $\bbN$. 
  We will adopt some temporary terminology and say that for $f \in F$, $\prod_{i=1}^m \bigl( a(i) \prod_{t \in H(i)} g_f(t) \bigr) a(m+1)$ has \emph{proper representation} if and only if $\prod_{i=1}^m \bigl( a(i) \prod_{t \in H(i)} g_f(t) \bigr) a(m+1) = x(n, c, t, f)$ where $c \in S^{n+1}$ is defined as follows:
  \[
    c(j) = 
    \begin{cases}
      a(j) & \mbox{if $j = 1$;} \\
      b & \mbox{if $s \in \{0, 1, \ldots, m-1\}$ and} \\
      &   \hspace{2em}\mbox{$2+h_s \le j \le h_s$; and} \\
      ba(s+1) & \mbox{if $s \in \{1, 2, \ldots, m\}$ and $j = 1+h_s$.}
    \end{cases}
  \]
  {
    (To see how our $c$ was derived let's consider a reasonable small example with $m = 3$, $H(1) = \{3, 5\}$, $H(2) = \{7\}$, and $H(3) = \{9, 11, 15\}$.
    With these numbers we have
    \begin{align*}
      \textstyle
      \prod_{i=1}^3 \bigl( a(i) \prod_{t \in H(i)} g(t) \bigr) a(4) &=
      a(1) g(3)g(5) a(2) g(7) a(3) g(9)g(11)g(15) a(4) \\
      &= a(1) f(3)bf(5)b a(2) f(7)b a(3) f(9)bf(11)bf(15)b a(4).
    \end{align*}
    Therefore $n = 6$, $c(1) = a(1)$, $c(2) = b$, $c(3) = ba(2)$, $c(4) = ba(3)$, $c(5) = b$,  $c(6) = b$, and $c(7) = ba(4)$.)
  }

  We prove that for $f \in F$, $\prod_{i=1}^m \bigl( a(i) \prod_{t \in H(i)} g_f(t) \bigr) a(m+1)$ has proper representation by induction on $m$. 
  First suppose that $m = 1$, then 
  \[
    \textstyle
    \prod_{i=1}^1 \bigl( a(i) \prod_{t \in H(i)} g_f(t) \bigr) a(2) = a(1) f(t_1)b  f(t_2)b \cdots f(t_n)b a(2).
  \]
  In this case $h_0 = 0$, $h_1 = n$, and so $s$ can only be 0 or 1.
  If $2+h_0 = 2 \le j \le h_1 = n$, then by definition of $c$ we have $c(j) = b$ for all $j \in \{2, 3, \ldots, n\}$. 
  Also since $1 + h_1 = n+1$, we have $c(n+1) = ba(2)$. 
  Therefore
  \[
    \textstyle
    a(1) f(t_1)b  f(t_2)b \cdots f(t_n)b a(2) = c(1) f(t_1) c(2) f(t_2) c(3) \cdots f(t_n) c(n+1),
  \]
  and so $\prod_{i=1}^1 \bigl( a(i) \prod_{t \in H(i)} g_f(t) \bigr) a(2)$ has proper representation. 

  Now let $m > 1$ and assume that $\prod_{i=1}^{m-1} \bigl( a(i) \prod_{t \in H(i)} g_f(t) \bigr) a(m)$ has proper representation with $\prod_{i=1}^{m-1} \bigl( a(i) \prod_{t \in H(i)} g_f(t) \bigr) a(m) = x(n, c, t, f)$. 
  Then we have 
  \begin{align*}
    \textstyle
    \prod_{i=1}^{m} \bigl( a(i) \prod_{t \in H(i)} g_f(t) \bigr) a(m+1) &= \textstyle
    \prod_{i=1}^{m-1} \bigl( a(i) \prod_{t \in H(i)} g_f(t) \bigr) a(m) \prod_{t \in H(m)} g_f(t) a(m+1), \\
    &= \textstyle 
    x(n, c, t, f) \prod_{t \in H(m)} g_f(t) a(m+1).
  \end{align*}
  Now $c(n+1) \prod_{t \in H(m)} g_f(t) a(m+1)$ has proper representation, say with, 
  \[
    \textstyle
    \prod_{t \in H(m)} g_f(t) a(m+1) = x(p, d, u, f).
  \]
  By translating the indices for $u$ and $d$ it follows that 
  \[
    \textstyle
    \prod_{i=1}^{m} \bigl( a(i) \prod_{t \in H(i)} g_f(t) \bigr) a(m+1)
  \]
  has proper representation.
\end{proof}

\begin{thm}
  Let $(S, \cdot)$ be a semigroup and $A$ a subset of $S$.
  The following are equivalent. 
  \begin{itemize}
    \item[(a)] $A$ is a $J$-set.
    \item[(b)] For all $F \in \calT$, there exist $m \in \bbN$, $a \in S^{m+1}$, and $H \in \calI_m$ such that for all $f \in F$ we have $\prod_{i=1}^n\bigl( a(i) \prod_{t \in H(i)} f(t) \bigr) a(m+1) \in A$.
  \end{itemize}
\end{thm}
\begin{proof}
  \textsl{(a) $\Rightarrow$ (b)}
  Let $F \in \calT$ and pick $m \in \bbN$, $a \in S^{m+1}$, and $t \in \calJ_m$ as guaranteed. 
  For each $i \in \{1, 2, \ldots, m\}$ put $H(i) = \{ t_i\}$. 
  Then $H \in \calI_m$ and the conclusion follows. 

  \textsl{(b) $\Rightarrow$ (a)}
  For all $F \in \calT$ by Lemma \ref{lem:rewrite-jset} there exist $m \in \bbN$, $a \in S^{m+1}$, and $t \in \calJ_m$ such that for all $f \in F$, $x(m, a, t, f) \in A$.
  Therefore $A$ is a $J$-set.
\end{proof}

With this theorem all of the results previously proved about $J$-sets apply  with our definition.
However in this chapter, we choose to reprove these results primarily to show how the proofs can be simplified with the ``new'' definition of $J$-set.

The first result we (re)prove is \cite[Lemma 2.4]{Hindman:2010fk} which shows that when the underlying semigroup is commutative the definition of a $J$-set is considerably simplified. 

\begin{lem}
  \label{lem:comm-jsets}
  Let $(S, +)$ be a commutative semigroup and let $A \subseteq S$. 
  Then $A$ is a $J$-set if and only if for every $F \in \calT$, there exist $a \in S$ and $H \in \Pf(\bbN)$ such that for all $f \in F$, $a + \sum_{t \in H} f(t) \in A$.
\end{lem}
\begin{proof}
  ($\Rightarrow$)
  Let $F \in \calT$.
  Pick $m \in \bbN$, $a \in S^{m+1}$, and $t \in \calJ_m$ such that $x(m, a, t, f) \in A$.
  Put $b = \sum_{i=1}^{m+1} a(i)$ and $H = \{t_1, t_2, \ldots, t_m\}$.
  Since $S$ is commutative we have that $b + \sum_{t \in H} f(t) = x(m, t, a, f) \in A$.

  ($\Leftarrow$)
  Let $F \in \Pf(\calT)$ and fix $b \in S$.
  For each $f \in F$ define $g_f \in \calT$ by $g_f(t) = f(t)+b$. 
  Pick $a \in S$ and $H \in \Pf(\bbN)$ such that for all $f \in F$, $a + \sum_{t \in H} g_f(t) \in A$. 
  Put $m = |H|$ and enumerate $H$ as a strictly increasing sequence $t_1 < t_2 < \cdots < t_m$. 
  Define $c \in S^{m+1}$ by $c(1) = a$ and $c(i) = b$ for every $i \in \{2, 3, \ldots, m+1\}$.
  Then $x(m, c, t, f) = a + \sum_{t \in H} g_f(t) \in A$.
\end{proof}

We next show that $J$-sets satisfy a superficially stronger condition. 
(In \cite[Definition 3.3(e)]{De:2008uq} this superficially stronger condition was used to define a $J$-set.)

\begin{lem}
  \label{lem:jset-start}
  Let $(S, \cdot)$ be a semigroup and $A$ a $J$-set in $S$.
  Then for every $F \in \Pf(\calT)$ and each $n \in \bbN$ there exist $m \in \bbN$, $a \in S^{m+1}$, and $t \in \calJ_m$, with $t_1 > n$ such that for all $f \in F$, $x(m, a, t, f) \in A$.
\end{lem}
\begin{proof}
  For each $f \in F$ define $g_f \in \calT$ by $g_f(t) = f(n+t)$.
  Since $A$ is a $J$-set, pick $m \in \bbN$, $a \in S^{m+1}$, and $t \in \calJ_m$ such that for all $f \in F$, $x(m, a, t, g_f) \in A$. 
  Define $u \in \calJ_m$ by $u_i =  n + t_i$.
  Then $x(m, a, u, f) = x(m, a, t, g_f) \in A$ for every $f \in F$.
\end{proof}

The next result we reprove is \cite[Theorem 3.4]{De:2008uq} which states that the set of all ultrafilters on a semigroup, each of whose member is a $J$-set, is a closed ideal.
To state this result we introduce the following notation.

\begin{defn}
  Let $(S, \cdot)$ be a semigroup. 
  Define
  \[
    J(S) = \{\, p \in \beta S : \mbox{for every $A \in p$, $A$ is a $J$-set} \,\}.
  \]
\end{defn}

\begin{thm}
  Let $(S, \cdot)$ be a semigroup.
  If $J(S)$ is nonempty, then $J(S)$ is a closed two-sided ideal of $\beta S$.
\end{thm}
\begin{proof}
  Let $p \not\in J(S)$ and pick $A \in p$ such that $A$ is not a $J$-set.
  By definition of $J(S)$ we must have $\overline{A} \cap J(S) = \emptyset$.
  Since $\overline{A}$ is a (basic) open neighborhood of $p$, it follows that $J(S)$ is topologically closed in $\beta S$.

  Now let $p \in J(S)$ and $q \in \beta S$.
  To see that $J(S)$ is an ideal, we show that $pq \in J(S)$ and $qp \in J(S)$. 

  We first show that $pq \in J(S)$.
  Let $F \in \Pf(\calT)$, let $A \in pq$, and put $B = \{\, x \in S : x^{-1}A \in q \,\}$.
  Then $B \in p$ and so $B$ is a $J$-set.
  Pick $m \in \bbN$, $a \in S^{m+1}$, and $t \in \calJ_m$ such that for all $f \in F$ we have $x(m, a, t, f) \in B$.
  By definition of $B$ this means that for all $f \in F$, $x(m, a, t, f)^{-1}A \in q$. 
  Since $q$ is an ultrafilter and $F$ is finite, we have $\bigcap_{f \in F} x(m, a, t, f)^{-1}A \in q$, and therefore, $\bigcap_{f \in F} x(m, a, t, f)^{-1}A \ne \emptyset$.
  Pick $b \in \bigcap_{f \in F} x(m, a, t, f)^{-1}A$ and define $c \in S^{m+1}$ by
  \[
    c =
    \begin{cases}
      a(j) & \mbox{if $j \in \{1, 2, \ldots, m\}$,} \\
      a(m+1)b & \mbox{if $j = m+1$.}
    \end{cases}
  \]
  Therefore, for each $f \in F$, $x(m, c, t, f) \in A$, that is, $A$ is a $J$-set and so $pq \in J(S)$.

  Now we show that $qp \in J(S)$.
  Again let $F \in \Pf(\calT)$, let $A \in qp$, and put $B = \{\, x \in S : x^{-1}A \in p \,\}$.
  Then $B \in q$, and therefore $B \ne \emptyset$.
  Pick $b \in B$, then $b^{-1}A \in p$ and so $b^{-1}A$ is a $J$-set.
  Pick $m \in \bbN$, $a \in S^{m+1}$, and $t \in \calJ_m$ such that for all $f \in F$, we have $x(m, a, t, f) \in b^{-1}A$.
  Define $c \in S^{m+1}$ by
  \[
    c =
    \begin{cases}
      ba(1) & \mbox{if $j =1$,} \\
      a(j) & \mbox{if $j \in \{2, \ldots, m, m+1\}$.}
    \end{cases}
  \]
  Therefore, for each $f \in F$, $x(m, c, t, f) \in A$, that is, $A$ is a $J$-set and so $qp \in J(S)$.
\end{proof}

Even though it is clear that every semigroup $S$ is a $J$-set in itself, this easy fact doesn't necessarily imply that $J(S) \ne \emptyset$.
It's true that $J(S)$ \emph{is} nonempty but this nontrivial fact follows either from the fact that piecewise syndetic sets are $J$-sets, Lemma \ref{lem:pr-jsets} and \cite[Theorem 3.11]{Hindman:1998fk}, or the Central Sets Theorem. 

We start by proving that piecewise syndetic sets are $J$-sets.
The following result is modeled after \cite[Theorems 14.1 and 14.7]{Hindman:1998fk}. 

\begin{thm}
  Let $(S, \cdot)$ be a semigroup and $A \subseteq S$ piecewise syndetic.
  Then $A$ is also a $J$-set.
\end{thm}
\begin{proof}
  Let $F \in \calT$, put $k = |F|$, and enumerate $F$ as $\{f_1, f_2, \ldots, f_k\}$.
  Put $Y = \bigtimes_{t=1}^k \beta S$, then with the product topology $Y$ is a compact right-topological semigroup by \cite[Theorem 2.22]{Hindman:1998fk}
  For each $i \in \bbN$ define
  \begin{align*}
    I_i &= \Bigl\{\, \bigl( x(m, a, t, f_1), x(m, a, t, f_2), \ldots,
    x(m, a, t, f_k) \bigr) : \mbox{$m \in \bbN$, $a \in S^{m+1}$,} \\
    &\hspace{5em} \mbox{$t \in \calJ_m$, and $t_1 > i$}
    \,\Bigr\},
  \end{align*}
  and put $E_i = I_i \cup \{\, (a, a, \ldots, a) : a \in S \,\}$.
  Put $I = \bigcap_{i=1}^\infty \overline{I_i}$ and $E = \bigcap_{i=1}^\infty \overline{E_i}$.

  Observe that $I$ and $E$ are nonempty closed subsets of $Y$.
  (Since each $I_i \ne \emptyset$ and $I_{i+1} \subseteq I_i$.)

  We claim that $E$ is a subsemigroup of $Y$ and $I$ is an ideal of $E$.
  Let $p$, $q \in E$, $U$ be an open neighborhood of $pq$, and let $i \in \bbN$. 
  Since $\rho_q$ is continuous, pick $V$ a neighborhood of $p$ such that $Vq \subseteq U$. 
  If $p \in I$, then pick $\vec{x} \in I_i \cap V$ otherwise pick $\vec{x} \in E_i \cap V$.
  If $\vec{x} \in I_i$, then pick $m \in \bbN$, $a \in S^{m+1}$, and $H \in \calI_{m,k}$ with $\min H(1) > i$ such that
  \[
    \vec{x} = \bigl( x(m, a, H, f_1), x(m, a, H, f_2), \ldots, x(m,
    a, H, f_l) \bigr)
  \]
  In this case put $j = \max H(m)$, otherwise put $j=i$. 

  Since $\lambda_{\vec{x}}$ is continuous, pick $W$ a neighborhood of $q$ such that $xW \subseteq U$. 
  If $q \in I$, then pick $\vec{y} \in I_j \cap W$, otherwise pick $\vec{y} \in E_j \cap W$.
  Then $\vec{x} \vec{y} \in E_i \cap U$, and if $p \in I$ or $q \in I$, then $\vec{x} \vec{y} \in I_i \cap U$. 
  Hence, it follows that $E$ is a subsemigroup of $Y$ and $I$ is an ideal of $E$.

  Pick $p \in K(\beta S) \cap \overline{A}$ and put $\overline{p} = (p, p, \ldots, p)$. 
  Since $K(Y) = \bigtimes_{t=1}^l K(\beta S)$, we have that $\overline{p} \in K(Y)$. 
  We show that $\overline{p} \in E$.
  Let $U$ be a neighborhood of $\overline{p}$ and pick $B_1$, $B_2$, \dots, $B_l \in p$ such that $\bigtimes_{t=1}^l \overline{B_t} \subseteq U$. 
  Pick $a \in \bigcap_{t=1}^l B_t$, then $(a, a, \ldots, a) \in U \cap E_i$ for all $i \in \bbN$. 
  Hence $\overline{p} \in E$ and moreover $\overline{p} \in K(Y) \cap E$.

  Since $K(Y) \cap E \ne \emptyset$, by Theorem \ref{thm:smallest-subsemigrp} we have $K(E) = K(Y) \cap E$.
  Therefore $\overline{p} \in K(E) \subseteq I$.
  Hence $I_i \cap \bigtimes_{i=1}^l A \ne \emptyset$ for all $i \in \bbN$ and so our conclusion follows.
\end{proof}

\begin{cor}
  Let $(S, \cdot)$ be a semigroup, then $c\ell\bigl( K(\beta S) \bigr) \subseteq J(S)$.
\end{cor}
\begin{proof}
  Let $p \in c\ell\bigl( K(\beta S) \bigr)$ and $A \in p$.
  By \cite[Corollary 4.41]{Hindman:1998fk}, $A$ is a piecewise syndetic set and so $A$ is a $J$-set.
  Therefore $p \in J(S)$.
\end{proof}

\begin{cor}
  If $A \subseteq \bbN$ be a piecewise syndetic set, then $A$ is also an $AP$-set. 
\end{cor}
\begin{proof}
  Since $A$ is also a $J$-set, it suffices to show that all $J$-sets in $\bbN$ are $AP$-set. 
  Let $k$ be a positive integer and let $\la x_n \ra_{n=1}^\infty$ be a sequence in $\bbN$. 
  For each $i \in \{1, 2, \ldots, k\}$ put $f_i(t) = i \cdot x_t$.
  Since $A$ is a $J$-set, pick $a \in \bbN$ and $H \in \Pf(\bbN$ such that for all $i \in \{1, 2, \ldots, k\}$, $a + \sum_{t \in H} f_i(t) \in A$. 
  Put $d = \sum_{t \in H} x_t$ and observe that for all $i \in \{1, 2, \ldots, k\}$ we have
  \begin{align*}
    a+id &= \textstyle a+ i(\sum_{t \in H} x_t) = a + \sum_{t \in H} ix_t, \\
    &= \textstyle a + \sum_{t \in H} f_i(t) \in A. 
  \end{align*}
  Hence $A$ is an $AP$-set.
\end{proof}

We now provide another proof, that when combined with \cite[Theorem 3.11]{Hindman:1998fk}, proves that $J(S)$ is nonempty.

\begin{lem}
  \label{lem:pr-jsets}
  Let $(S, \cdot)$ be a semigroup with $A_1 \subseteq S$ and $A_2 \subseteq S$.
  If $A_1 \cup A_2$ is a $J$-set, then either $A_1$ is a $J$-set or $A_2$ is a $J$-set.
\end{lem}
\begin{proof}
  Suppose, to the contrary, that both $A_1$ and $A_2$ are \emph{not} $J$-sets.
  Pick $F_1$ and $F_2$ in $\Pf(\calT)$ such that for all $m \in \bbN$, every $a \in S^{m+1}$, and every $t \in \calJ_m$, there exist $f \in F_1$ and $g \in F_2$ such that $x(m, a, t, f) \not\in A_1$ and $x(m, a, t, g) \not\in A_2$.
  Let $F = F_1 \cup F_2$, put $k = |F|$, and enumerate $F$ as $\{f_1, f_2, \ldots, f_k\}$.

  By Theorem \ref{thm:finitary-shj} pick $n \in \bbN$ such that whenever $\{1, 2, \ldots, k\}^n$ is 2-colored there exists a variable word $w(\star)$ which begins and ends with a constant letter, with no two adjacent variable words, and for which the combinatorial line $\bigl\{\, w(\ell) : \ell \{1, 2, \ldots, k\} \,\bigr\}$ is monochromatic. 
  (For instance, if $k = 4$ and $n = 10$, then $w(\star) = (1, \star, 1, 4, \star, 2, \star, 3, 3, 1)$ is a suitable variable word.)

  Now for each $w = (x_1, x_2, \ldots, x_n) \in \{1, 2, \ldots, k\}^n$ define $g_w \in \calT$ by $g_w(t) = \prod_{i=1}^n f_{x_i}(nt + i)$.
  We have, by hypothesis, that $A_1 \cup A_2$ is a $J$-set and since $\{1, 2, \ldots, k\}^n$ is finite, we may pick $m \in \bbN$, $a \in S^{m+1}$, and $t \in \calJ_m$ such that for all $w \in \{1, 2, \ldots, k\}^n$, $x(m, a, t, g_w) \in A$.

  Define the function $\varphi \colon \{1, 2, \ldots, k\}^n \to \{1, 2\}$ as follows:
  \[
    \varphi(w) = 
    \begin{cases}
      1 & \mbox{if $x(m, a, t, g_w) \in A_1$, and} \\
      2 & \mbox{otherwise.}
    \end{cases}
  \]
  Since $\varphi$ is a 2-coloring of $\{1, 2, \ldots, k\}^n$ we may apply the Hales-Jewett Theorem to pick a variable word $w(\star)$ which begins and ends with a constant letter, with no two adjacent variable letters, and for which the combinatorial line $\bigl\{\, w(\ell) : \ell \in \{1, 2, \ldots, k\} \,\bigr\}$ is monochromatic. 
  Without loss of generality we may suppose that $\varphi \bigl( w(\ell) \bigr) = 1$ for every $\ell \in \{1, 2, \ldots, k\}$.

  We claim that there exist $r \in \bbN$, $c \in S^{mr+1}$, and $s \in \calJ_{mr}$ such that for all $\ell \in \{1, 2, \ldots, k\}$ we have $x(mr, c, s, f_\ell) = x(m, a, t, g_{w(\ell)}) \in A_1$. 
  Observe that if we prove our claim, then we will have arrived at a contradiction. 
  Therefore for the rest of the proof we focus only on rewriting $x(m, a, t, g_{w(\ell)})$ into the appropriate form.

  Let $w(\star) = (x_1, x_2, \ldots, x_n)$ with each $x_i \in \{1, 2, \ldots, k\} \cup \{\star\}$. 
  Let $r$ be the number of variable letters in $w(\star)$, and for each $i \in \{1, 2, \ldots, r-1\}$, let $b_i$ be the position of the \mbox{$i$th} variable letter in $w(\star)$ and let $b_0$ be the position of the \mbox{$r$th} variable letter.
  (For instance, if $k = 4$ and $n = 10$, $w(\star) = (1, \star, 1, 4, \star, 2, \star, 3, 3, 1)$, then $r = 3$, $b_1 = 2$, $b_2 = 5$, and $b_0 = 7$.
  The reason we let $b_0$ represent the position of the \mbox{$r$th} variable letter is because we will be performing modulo arithmetic on the index of the $b_i$'s.)

  For each $i \in \{1, 2, \ldots, mr\}$ define $s_i = nt_{\lceil i/r \rceil} + b_{i \bmod r}$.
  Observe that $r \le n/2$, and since $t_1 < t_2 < \cdots < t_m$, it follows that $s = (s_1, s_2, \ldots, s_{mr}) \in \calJ_{mr}$.

  To define $c \in S^{mr+1}$ we will need to define some further helpful notation. 
  Let $L = \bigl( L(1), L(2), \ldots, L(r+1) \bigr)$ be a partition of $\{1, 2, \ldots, n\} \setminus \{b_1, b_2, \ldots, b_0\}$ such that each $L(i)$ is nonempty and $\max L(i) < \min L(i+1)$ for every $i \in \{1, 2, \ldots, r\}$. 
  For arithmetical convenience put $L(0) = L(r)$. 
  (For instance, if $k = 4$, $n = 10$, $w(\star) = (1, \star, 1, 4, \star, 2, \star, 3, 3, 1)$, then $L(1) = \{1\}$, $L(2) = \{3,4\}$, $L(3) = \{6\}$, $L(4) = \{8, 9, 10\}$, and $L(0) = \{6\}$.)

  Now we define $c \in S^{mr+1}$ as follows:
  \[
    c(i) = 
    \begin{cases}
      a(1)\prod_{j \in L(1)} f_j(nt_1 + j) & \mbox{if $i = 1$,} \\
      \prod_{j \in L(i \bmod r)}  f_j(nt_{\lceil i/(r+1) \rceil} + j) & \mbox{if $i \not\equiv 1 \pmod r$, }\\
      \prod_{j \in L(r+1)} \bigl( f_j(nt_{\lceil i/(r+1) \rceil} + j) \bigr) a\bigl( \lceil (i+1)/(r+1) \rceil \bigr) & \\
      \hspace{2em} \cdot\prod_{j \in L(1)} f_j(nt_{\lceil (i+1)/(r+1) \rceil} + j) & \mbox{if $i \bmod r \equiv 1$ and $i \ne mr +1$, }\\
      \prod_{j \in L(r+1)} \bigr( f_i(nt_m + j) \bigr) a(m+1) & \mbox{if $i = mr + 1$.}
    \end{cases}
  \]

  Then we have that $x(mr, c, s, f_\ell) = x(m, a, t, g_{w(\ell)})$ for all $\ell \in \{1, 2, \ldots, k\}$.
\end{proof}

With Lemma \ref{lem:pr-jsets} we prove a stronger relationship between $J$-sets and $J(S)$.

\begin{thm}
  \label{thm:jsets-ideal}
  Let $(S, \cdot)$ be a semigroup and $A \subseteq S$.
  Then $J(S) \cap \overline{A} \ne \emptyset$ if and only if $A$ is a $J$-set.
\end{thm}
\begin{proof}
  ($\Rightarrow$)
  Let $p \in J(S) \cap \overline{A}$. 
  Since $p \in J(S)$, every member of $p$ is a $J$-set.
  Hence $A$ is a $J$-set.

  ($\Leftarrow$)
  Suppose $A$ is a $J$-set. 
  By Lemma \ref{lem:pr-jsets} and \cite[Theorem 3.11]{Hindman:1998fk}, there exists an ultrafilter $p \in \beta S$ such that $A \in p$ and every member of $p$ is a $J$-set. 
  Therefore $p \in J(S)$.
\end{proof}


\section{Central Sets Theorem}
The final proof that $J(S)$ is nonempty follows from the Central Sets Theorem.
\begin{defn}
  Let $(S, \cdot)$ be a semigroup and $A \subseteq S$.
  We call $A$ a \textsl{$C$-set} if and only if there exist $m \colon \Pf(\calT) \to \bbN$, $\alpha \in \bigtimes_{F \in \Pf(\calT)} S^{m(F)+1}$, and $\tau \in \bigtimes_{F \in \Pf(\calT)} \calJ_{m(F)}$ such that 
  \begin{itemize}
    \item[(1)] if $F$, $G \in \Pf(\calT)$ with $F \subsetneq G$, then $\tau(F)\bigl( m(F) \bigr) < \tau(G)(1)$, and
    \item[(2)] whenever $n \in \bbN$, $G_1$,$G_2$, \dots, $G_n \in \Pf(\calT)$  with $G_1 \subsetneq G_2 \subsetneq \cdots \subsetneq G_n$, and for each $i \in \{1, 2, \ldots, n\}$, $f_i \in G_i$, then we have $\prod_{i=1}^n x(m(G_i), \alpha(G_i), \tau(G_i), f_i) \in A$.
  \end{itemize}
\end{defn}

We now prove that $C$-sets and idempotents in $J(S)$ are closely related.
\begin{thm}
  \label{thm:csets}
  Let $(S, \cdot)$ be a semigroup and let $A \subseteq S$. 
  Then $A$ is a $C$-set if and only if there exists an idempotent in $p \in \overline{A} \cap J(S)$.
\end{thm}
\begin{proof}
  ($\Rightarrow$)
  Pick $m \colon \Pf(\calT) \to \bbN$, $\alpha \in \bigtimes_{F \in \Pf(\calT)} S^{m(F)+1}$, and $\tau \in \bigtimes_{F \in \Pf(\calT)} \calJ_{m(F)}$ as guaranteed by the definition of a $C$-set. 
  For each $F \in \Pf(\calT)$ define 
  \begin{align*}
    T(F) &= \bigl\{\, \textstyle \prod_{i=1}^n x(m(F_i), \alpha(F_i), \tau(F_i), f_i) : \mbox{$n \in \bbN$, each $F_i \in \calT$,}\\
 &\hspace{12em}\mbox{$F \subsetneq F_1 \subsetneq F_2 \subsetneq \cdots \subsetneq F_n$, $\la f_i \ra_{i=1}^n \in \bigtimes_{i=1}^n F_i$,} \,\bigr\}.
  \end{align*}

  Observe that the collection $\{\, T(F) : F \in \Pf(\calT) \,\}$ has the finite intersection property since $T(F \cup G) \subseteq T(F) \cap T(G)$ for all $F$, $G \in \Pf(\calT)$.
  Therefore $\mathbf{Q} = \bigcap_{F \in \Pf(\calT)} c\ell_{\beta S}\bigl(T(F)\bigr)$ is a closed nonempty subset of $\beta S$. 

  We show that $\mathbf{Q}$ is in fact a subsemigroup of $\beta S$ contained in $c\ell_{\beta S} (A)$.
  To see that $\mathbf{Q}$ is a subsemigroup it suffices, by \cite[Theorem 4.20]{Hindman:1998fk}, to show that for all $F \in \Pf(\calT)$ and for every $y \in T(F)$, there exists $G \in \Pf(\calT)$ such that $yT(G) \subseteq T(F)$. 
  So let $F \in \Pf(\calT)$ and $y \in T(F)$.
  Pick $n \in \bbN$, for every $i \in \{1, 2, \ldots, n\}$ pick $F_i \in \calT$ such that $F \subsetneq F_1 \subsetneq F_2 \subsetneq \cdots F_n$, and $\la f_i \ra_{i=1}^n \in \bigtimes_{i=1}^n F_i$ such that $y = \prod_{i=1}^n x(m(F_i), \alpha(F_i), \tau(F_i), f_i)$. 
  We prove that we can choose $G = F_n$.
  Let $z \in T(F_n)$ and pick $m \in \bbN$, for every $i \in \{1, 2, \ldots, m\}$ pick $G_i \in \calT$ such that $F_n \subsetneq G_1 \subsetneq G_2 \subsetneq \cdots \subsetneq G_m$, and pick $\la g_i \ra_{i=1}^m \in \bigtimes_{i=1}^m G_i$ such that $z = \prod_{i=1}^m x(m(G_i), \alpha(G_i), \tau(G_i), g_i)$.

  For each $i \in \{1, 2, \ldots, n+m\}$ define 
  \[
    H_i = 
    \begin{cases}
      F_i & \mbox{if $i \in \{1, 2, \ldots, n\}$,} \\
      G_{i-n} & \mbox{if $i \in \{n+1, n+2, \ldots, n+m\}$,}
    \end{cases}
  \]
  and define the sequence
  \[
    h_i =
    \begin{cases}
      f_i & \mbox{if $i \in \{1, 2, \ldots, n\}$,} \\
      g_{i-n} & \mbox{if $i \in \{n+1, n+2, \ldots, n+m\}$.}
    \end{cases}
  \]
  Then $F \subsetneq H_1 \subsetneq H_2 \subsetneq \cdots subsetneq H_{n+m}$, $\la h_i \ra_{i=1}^{n+m}$, and $yz = \prod_{i=1}^{n+m} x(m(H_i), \alpha(H_i), \tau(H_i), h_i) \in T(F)$. 
  Hence $\mathbf{Q}$ is a subsemigroup of $\beta S$. 

  By definition, $T(F) \subseteq A$ for every $F \in \Pf(\calT)$, and so $c\ell_{\beta S} (T(F)) \subseteq c\ell_{\beta S} (A)$. 
  Hence $\mathbf{Q} \subseteq c\ell_{\beta S} (A)$.

  We now claim that $K(\mathbf{Q}) \subseteq \overline{A} \cap J(S)$. 
  If this claim is true, then we are done since any idempotent in $K(\mathbf{Q})$ will establish the theorem. 
  Since $\mathbf{Q} \subseteq c\ell_{\beta S} (A)$, it evident that $\mathbf{Q} \subseteq c\ell_{\beta S} (A)$. 

  Now let $p \in K(\mathbf{Q})$ and $B \in p$.
  Using \cite[Lemma 14.9]{Hindman:1998fk} we shall show that $B$ is a $J$-set. 
  Let $F \in \Pf(\calT)$ and put $D = \{\, G \in \Pf(\calT) : F \subseteq G \,\}$. 
  Observe that $(D, \subseteq)$ is a directed set and $\mathbf{Q} = \bigcap_{G \in D} c\ell_{\beta S} (T(G))$.

  For $G \in D$, define $I(G) \subseteq \bigtimes_{f \in F} S$ as follows: for $w \in \bigtimes_{f \in F} S$, $w \in I(G)$ if and only if there is some $n \in \bbN \setminus \{1\}$ such that 
  \begin{itemize}
    \item[(1)]
      there exist disjoint nonempty sets $C_1$ and $C_2$ with $\{1, 2, \ldots, n\} = C_1 \cup C_2$, 
      
    \item[(2)]
      there exist a strictly increasing sequence $\la G_i \ra_{i=1}^n$ in $\Pf(\calT)$ with $G \subsetneq G_1$, and

    \item[(3)]
      there exists $\sigma \in \bigtimes_{i \in C_1} G_i$, 
  \end{itemize}
such that for every $f \in F$, if $\gamma_f \in \bigtimes_{i=1}^n G_i$ is defined by 
  \[
    \gamma_f(i) = 
    \begin{cases}
      \sigma(i) & \mbox{if $i \in C_1$,} \\
      f & \mbox{if $i \in C_2$,} 
    \end{cases}
  \]
then $w(f) = \prod_{i=1}^n x(m(G_i), \alpha(G_i), \tau(G_i), \gamma_f(i))$.
  For each $G \in D$, define $E(G) = I(G) \cup \{\, \overline{b} : b \in T(G) \,\}$. 

  We claim that $\la E(G) \ra_{G \in D}$ and $\la I(G) \ra_{G \in D}$ satisfy statements (a), (b), (c), and (e) of \cite[Lemma 14.9]{Hindman:1998fk}.

  Assume, temporarily, that our claim is true. 
  Then by \cite[Lemma 14.9]{Hindman:1998fk} if $Y = \bigtimes_{f \in F} \beta S$, $E = \bigcap_{G \in D} c\ell_Y(E(G))$, and $I = \bigcap_{G \in D} c\ell_Y(I(G))$, then $E$ is a subsemigroup of $Y$, $I$ is an ideal of $E$, and for every $p \in K(\mathbf{Q})$, $\overline{p} \in K(E) \subseteq I$. 
  Since $\bigtimes_{f \in F} c\ell_{\beta S} (B)$ is a neighborhood of $\overline{p}$ we can pick $w \in I(F) \cap \bigtimes_{f \in F} c\ell_{\beta S} (B)$. 
  Pick $n \in \bbN\setminus \{1\}$, $C_1$, $C_2$, $\la G_i \ra_{i=1}^n$, and $\sigma \in \bigtimes_{i \in C_1} G_i$ as guaranteed by the definition of $I(F)$. 
  Put $r = |C_2|$ and enumerate $C_2$ as a strictly increasing sequence $h_1$, $h_2$, \dots, $h_r$.
  Put $u = \sum_{i=1}^r m(G_i)$.
  We define $c \in S^{u+1}$ and $t \in \calJ_u$ such that for all $f \in F$, $x(u, c, t, f) \in B$.

  Define $c \in S^{u+1}$ as follows:
  \[
    c(1) = 
    \begin{cases}
      \alpha(G_1)(1) & \mbox{if $h_1 = 1$,} \\
      \bigl(\prod_{i=1}^{h_1-1} x(m(G_i), \alpha(G_i), \tau(G_i), \sigma(i))\bigr) \cdot \alpha(G_{h_1})(1) & \mbox{if $h_1 > 1$.}
    \end{cases}
  \]
  For each positive integer $j$ with $1 < j < m(G_{h_1})$ put $c(j) = \alpha(G_{h_2})(j)$, and for each positive integer $j$ with $1 \le j \le m(G_{h_1})$ define $t_j = \tau(G_{h_1})(1)$. 

  Now for each $s \in \{1, 2, \ldots, u-1\}$ put $v_s = \sum_{i=1}^s m(G_{h_i})$, and define
  \[
    c(v_s+1) = 
    \begin{cases}
      \alpha(G_{h_s})(m(G_{h_s}+1))\alpha(G_{h_{s+1}})(1) & \mbox{if $h_{s+1} = h_s + 1$,} \\

      \alpha(G_{h_s})(m(G_{h_s}+1))\cdot & \\
      \hspace{3.5em}\bigl(\prod_{i=h_{s+1}}^{h_{s+1}-1} x(m(G_i), \alpha(G_i), \tau(G_i), \sigma(i))\bigr)\alpha(G_{h_{s+1}})(1) & \mbox{if $h_{s+1} > h_s+1$.}
    \end{cases}
  \]
  For each $s \in \{1, 2, \ldots, u-1\}$ and every positive integer $j$ with $v_s < j \le \sum_{i=1}^{s+1} m(G_{h_i})$ put $t_j = \tau(G_{h_{s+1}})(j-u)$. 
  Finally, we define
  \[
    c(u+1) = 
    \begin{cases}
      \alpha(G_{h_r})(m(G_n) + 1) & \mbox{if $h_r = n$,} \\
      \alpha(G_{h_r})(m(G_{h_r}+1)\prod_{i=h_r+1}^n x(m(G_i), \alpha(G_i), \tau(G_i), \sigma(i)) & \mbox{if $h_r < n$.}
    \end{cases}
  \]

  Then for every $f \in F$, $x(u, c, t, f) \in B$, and so $B$ is a $J$-set.

  % Put in proof of claim.
  ($\Leftarrow$)
  Let $p = p \cdot p \in \overline{A} \cap J(S)$.
  Since $A \in p$ and $p$ is an idempotent we have, by \cite[Lemma 4.14]{Hindman:1998fk}, that $x^{-1}A^\star \in p$ for all $x \in A^\star$ where $A^\star = \{\, x \in A : x^{-1}A \in p\}$. 
  We will recursively define our functions $m$, $\alpha$, and $\tau$ by the size of $F \in \Pf(\calT)$ such that we satisfy, for $F \in \Pf(\calT)$, the following hypothesis:
  \begin{itemize}
    \item[(1)] If $\emptyset \ne G \subsetneq F$, then $\tau(G)\bigl( m(G) \bigr) < \tau(F)(1)$.
    
    \item[(2)] If $n \in \bbN$, $\emptyset \ne G_1 \subsetneq G_2 \subsetneq \cdots \subsetneq G_n = F$, and $\la f_i \ra_{i=1}^n \in \bigtimes_{i=1}^n G_i$, then \[\textstyle \prod_{i=1}^n x(m(G_i), \alpha(G_i), \tau(G_i), f_i)) \in A^\star.\]
  \end{itemize}

  Let $F \in \Pf(\calT)$.
  First, assume that $|F| = 1$, that is $F = \{f\}$ for some sequence $f \in \calT$.
  Since $A^\star$ is a $J$-set, pick $m(F) \in \bbN$, $\alpha(F) \in S^{m(F)+1}$, and $\tau(F) \in \calJ_{m(F)}$ such that $x(m(F), \alpha(F), \tau(F), f) \in A^\star$. 

  Now assume that $|F| > 1$ and for all $\emptyset \ne G \subsetneq F$ we have defined $m(G)$, $\alpha(G)$, and $\tau(G)$ so that hypotheses (1) and (2) hold.
  Put 
  \begin{align*}
    M &= \bigl\{\, \textstyle \prod_{i=1}^n x(m(G_i), \alpha(G_i), \tau(G_i), f_i) : \mbox{$n \in \bbN$, $\emptyset \ne G_1 \subsetneq G_2 \subsetneq \cdots \subsetneq G_n \subsetneq F$} \\
    &\hspace{12em} \mbox{and $\la f_i \ra_{i=1}^n \in \bigtimes_{i=1}^n G_i$} \,\bigr\}.
  \end{align*}
  Observe that since $F$ is finite, $M$ is also finite.
  By hypothesis (2) we have that $M \subseteq A^\star$.
  Put $B = A^\star \cap \bigcap_{x \in M} x^{-1}A^\star$, then $B \in p$ and so $B$ is a $J$-set. 

  For each $\emptyset \ne G \subsetneq F$, put $l(G) = \tau(G)\bigl( m(G) \bigr)$ and put $k = \max\{\, l(G) : \emptyset \ne G \subsetneq F \,\}$.
  By Lemma \ref{lem:jset-start}, pick $m(F) \in \bbN$, $\alpha(F) \in S^{m(F)+1}$, and $\tau(F) \in \calJ_{m(F)}$ such that $\tau(F)(1) > k$ and for every $f \in F$, $x(m(F), \alpha(F), \tau(F), f) \in B$.
  Hypothesis (1) is satisfied, since $\tau(F)(1) > k \ge \max l(G)$ for all $\emptyset \ne G \subsetneq F$.
  We show that hypothesis (2) is also satisfied.
  Let $n \in \bbN$, $\emptyset \ne G_1 \subsetneq G_2 \subsetneq \cdots \subsetneq G_{n-1} \subsetneq G_n = F$, and $\la f_i \ra_{i=1}^n \bigtimes_{i=1}^n G_i$. 
  If $n = 1$, then $x(m(G_1), \alpha(G_1), \tau(G_1), f) \in B \subseteq A^\star$.
  Now assume that $n > 1$ and put $y = \prod_{i=1}^{n-1} x(m(G_i), \alpha(G_i), \tau(G_i), f_i)$. 
  By definition $y \in M$ and since $x(m(G_n), \alpha(G_n), \tau(G_n), f_n) = x(m(F), \alpha(F), \tau(F), f_n) \in B \subseteq y^{-1}A^\star$ we have
  \[
    \textstyle
    \prod_{i=1}^n x(m(G_i), \alpha(G_i), \tau(G_i), f_i) = y \cdot x(m(F), \alpha(F), \tau(F), f_n) \in A^\star.
  \]
  Hence hypotheses (1) and (2) are satisfied and this completes the proof for this direction.
\end{proof}

It's reasonably clear that any semigroup $S$ is a $C$-set in itself.
Then Theorem \ref{thm:csets} shows that there is an idempotent in $J(S)$ and hence $J(S)$ is nonempty.

\begin{cor}[Central Sets Theorem]
  Every central set in a semigroup is a $C$-set.
\end{cor}
\begin{proof}
  Let $(S, \cdot)$ be a semigroup with $A \subseteq S$ a central set. 
  Since $A$ is a central set, there exists an idempotent $p \in K(\beta S)$ with $A \in p$.
  Since $K(S) \subseteq J(S)$, we have that $A$ is also a $C$-set.
\end{proof}

Similar to how the definition of a $J$-set is simpler when the underlying semigroup is commutative we have the following result which shows that $C$-sets also have a simpler characterization when the underlying semigroup is commutative. 

\begin{lem}
  Let $(S, +)$ be a commutative semigroup and $A \subseteq S$.
  Then $A$ is a $C$-set if and only if there exist functions $\alpha \colon \Pf(\calT) \to S$ and $H \colon \Pf(\calT) \to \Pf(\bbN)$ such that
  \begin{itemize}
    \item[(1)] if $F$, $G \in \Pf(\calT)$ with $F \subsetneq G$, then $\max H(F) < \min H(G)$, and
    
    \item[(2)] whenever $m \in \bbN$, $G_1$, $G_2$, \dots, $G_m \in \Pf(\calT)$ with $G_1 \subsetneq G_2 \subsetneq \cdots \subsetneq G_m$, and for each $i \in \{1, 2, \ldots, m\}$, $f_i \in G_i$, we have $\sum_{i=1}^m\bigl( \alpha(G_i) + \sum_{t \in H(G_i)} f_i(t)\bigr) \in A$.
  \end{itemize}
\end{lem}
\begin{proof}
  ($\Rightarrow$)
  Since $A$ is a $C$-set, we also have that $A$ is a $J$-set in a commutative semigroup.
  By Theorem \ref{thm:csets} we also know that $A \in p$ for some idempotent $p \in J(S)$. 
  We will recursively define our functions $\alpha$ and $H$ by the size of $F \in \Pf(\calT)$ such that we satisfy, for $F \in \Pf(\calT)$, following hypotheses:
  \begin{itemize}
    \item[(i)] If $\emptyset \ne G \subsetneq F$, then $\max H(G) < \min H(F)$.
      
    \item[(ii)] If $n \in \bbN$, $\emptyset \ne G_1 \subsetneq G_2 \subsetneq \cdots \subsetneq G_n = F$, and $\la f_i \ra_{i=1}^n \in \bigtimes_{i=1}^n G_i$, then $\sum_{i=1}^n\bigl( \alpha(G_i) + \sum_{t \in H(G_i)} f_i(t)\bigr) \in A^\star$. 
  \end{itemize}
  
  Let $F \in \Pf(\calT)$.
  First, assume that $|F| = 1$, that is, $F = \{f\}$ for some sequence $f \in \calT$.
  Since $A^\star$ is a $J$-set (in a commutative semigroup), we may pick $\alpha(F) \in S$ and $H(F) \in \Pf(\bbN)$ such that $\alpha(F) + \sum_{t \in H(F)} f(t) \in A^\star$.

  Now assume that $|F| > 1$ and for all $\emptyset \ne G \subsetneq F$ we have defined $\alpha(G)$ and $H(G)$ so that hypotheses (i) and (ii) hold.
  Put
  \begin{align*}
    M &=\textstyle \bigl\{\, \sum_{i=1}^n \bigl( \alpha(G_i) + \sum_{t \in G_i} f_i(t) \bigr) : \mbox{$n \in \bbN$, $\emptyset \ne G_1 \subsetneq G_2 \subsetneq \cdots \subsetneq G_n \subsetneq F$,} \\
    &\hspace{12em} \mbox{and $\la f_i \ra_{i=1}^n \in \bigtimes_{i=1}^n G_i$} \,\bigr\}.
  \end{align*}
  Since $F$ is finite, $M$ is also finite and by hypothesis (ii) we have that $M \subseteq A^\star$.
  Put $B = A^\star \cap \bigcap_{x \in M} -x+A^\star$, then $B \in p$ and so $B$ is also a $J$-set.

  Put $k = \max \bigl( \bigcup\{\, H(G) : \emptyset \ne G \subsetneq F\,\} \bigr)$.
  Since $B$ is a $J$-set, pick $\alpha(F) \in S$ and $H(F) \in \Pf(\bbN)$ with $\min H(F) > k$ such that for all $f \in F$, $\alpha(F) + \sum_{t \in H(F)} f(t) \in B$. 

  Now hypothesis (i) is satisfied since $\min H(F) > k \ge \max H(G)$ for all $\emptyset \ne G \subsetneq F$. 
  We show that hypothesis (ii) is also satisfied.
  Let $n \in \bbN$, $\emptyset \ne G_1 \subsetneq G_2 \subsetneq \cdots \subsetneq G_n = F$, and $\la f_i \ra_{i=1}^n \in \bigtimes_{i=1}^n G_i$. 
  If $n = 1$, then $\sum_{i=1}^n \bigl( \alpha(G_i) + \sum_{t \in H(G_i)} f_i(t)\bigr) = \alpha(F) + \sum_{t \in H(F)} f_1(t) \in B \subseteq A^\star$. 
  Now assume that $n > 1$ and let $y = \sum_{i=1}^{n-1}\bigl( \alpha(G_i) + \sum_{t \in H(G_i)} f_i(t) \bigr)$.
  Then $y \in M$ and so $\alpha(G_n) + \sum_{t \in H(G_n)} f_n(t) = \alpha(F) + \sum_{t \in H(F)} f_n(t) \in B \subseteq -y + A^\star$. 
  Hence $\sum_{i=1}^n\bigl( \alpha(G_i) + \sum_{t \in H(G_i)} f_i(t) \bigr) \in A^\star$.

  ($\Leftarrow$)
  Pick $\alpha \colon \Pf(\calT) \to S$ and $H \colon \Pf(\calT) \to \Pf(\bbN)$ as guaranteed. 
  Observe that $A$ satisfies the necessity condition of Lemma \ref{lem:comm-jsets} and hence that $A$ is a $J$-set in $S$. 
  Therefore by Theorem \ref{thm:jsets-ideal} $\overline{A} \cap J(S) \ne \emptyset$.
  % Still need to provide more details.  Will probably have to use the big Lemma.

  Clearly $A$ is a $J$-set in a commutative semigroup and so we can rewrite the terms $\alpha(G_i) + \sum_{t \in H(G_i)} f_i(t)$ as $x(m(G_i), \beta(G_i), \tau(G_i), f_i)$. 
\end{proof}

% Endnotes 
% \theendnotes


\bibliographystyle{amsplain}
\bibliography{../references}

\end{document}